---
title: "ds2003 project"
author: "Philip Yao, Joseph Choe, Eunice Kim, Tu-Yen Dang, Herin Seo"
date: "2023-10-04"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r packages, message=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## Read in Data

- ```UnemploymentAndMentalIllnessSurvey.csv```
- ```MedicalStudentMentalHealth.csv```

```{r}
data1 <- read.csv("data/UnemploymentAndMentalIllnessSurvey.csv")
data2 <- read.csv("data/MedicalStudentMentalHealth.csv")
```



```{r}

#Clean Dataset 1
data1a <- data1%>%select(I.am.currently.employed.at.least.part.time, Education, Tiredness, Anxiety, Depression,)%>%mutate(Stress = Tiredness, Employed = I.am.currently.employed.at.least.part.time)
data1a <- data1a%>%select(Employed, Education, Stress, Anxiety, Depression,)
data1a[data1a == "Completed Undergraduate"] <- "Undergraduate"
data1a[data1a == "Completed Masters"] <- "Masters"
data1a[data1a == "Completed Phd"] <- "Phd"
data1a[data1a == "Some Phd"] <- "Masters"
data1a[data1a == "Some Undergraduate"] <- "High School"
data1a[data1a == "Some Masters"] <- "Undergraduate"
data1a[data1a == "Some highschool"] <- "Dropped out"
data1a[data1a == "High School or GED"] <- "High School"

#Clean Dataset 2
data2a <- data2%>%select(job,psyt,cesd,stai_t)%>%mutate(Education = "Undergraduate", Stress = psyt, Anxiety = ifelse(stai_t < 30,0,1), Depression = ifelse(cesd < 16, 0, 1), Employed = job)
data2a <- data2a%>%select(Employed, Education, Stress, Anxiety, Depression)
data2a <- data2a%>%mutate(Anxiety = as.integer(Anxiety), Depression = as.integer(Depression))
head(data2a)

#Merge Datasets
data3 <- bind_rows(data1a, data2a)
head(data3)
```
## Project Questions and Sources

1. How much does unemployment have an effect on stress, anxiety, and depression levels?

2. How does household income among those who are unemployed affect their mental health?

3. Does geographic location among those who are unemployed have an effect on their mental health?

4. How does education level among those who are unemployed affect their mental health?

- Data Sources:

[Unemployment and Mental Illness Survey](https://www.kaggle.com/datasets/michaelacorley/unemployment-and-mental-illness-survey)

[Medical Student Mental Health](https://www.kaggle.com/datasets/thedevastator/medical-student-mental-health?select=Data+Carrard+et+al.+2022+MedTeach.csv)
