---
title: "ds2003 project"
author: "Philip Yao, Joseph Choe, Eunice Kim, Tu-Yen Dang, Herin Seo"
date: "2023-10-04"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r packages, message=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## Files:

- ```UnemploymentAndMentalIllnessSurvey.csv```
- ```MedicalStudentMentalHealth.csv```

```{r}
data1 <- read.csv("data/UnemploymentAndMentalIllnessSurvey.csv")
data2 <- read.csv("data/MedicalStudentMentalHealth.csv")
```


## Data Cleaning and Merging:

```{r}

#Clean Dataset 1
data1a <- data1%>%select(I.am.currently.employed.at.least.part.time, Education, Tiredness, Anxiety, Depression,I.identify.as.having.a.mental.illness)%>%mutate(Stress = Tiredness, Employed = I.am.currently.employed.at.least.part.time, Mental_Illness_Status = I.identify.as.having.a.mental.illness)
data1a <- data1a%>%select(Employed, Education, Stress, Anxiety, Depression, Mental_Illness_Status)
data1a[data1a == "Completed Undergraduate"] <- "Undergraduate"
data1a[data1a == "Completed Masters"] <- "Masters"
data1a[data1a == "Completed Phd"] <- "PHD"
data1a[data1a == "Some Phd"] <- "Masters"
data1a[data1a == "Some Undergraduate"] <- "High School"
data1a[data1a == "Some Masters"] <- "Undergraduate"
data1a[data1a == "Some highschool"] <- "Dropped Out"
data1a[data1a == "High School or GED"] <- "High School"

#Clean Dataset 2
data2a <- data2%>%select(job,psyt,cesd,stai_t)%>%mutate(Education = "Undergraduate", Stress = psyt, Anxiety = ifelse(stai_t < 30,0,1), Depression = ifelse(cesd < 16, 0, 1), Employed = job)
data2a <- data2a%>%select(Employed, Education, Stress, Anxiety, Depression)
data2a <- data2a%>%mutate(Anxiety = as.integer(Anxiety), Depression = as.integer(Depression))
head(data2a)

#Merge Datasets
data3 <- bind_rows(data1a, data2a)
head(data3)
```
## Research Questions

Focus: Exploring The Intersection Between Unemployment and Mental Health

1. How much does unemployment have an effect on stress, anxiety, and depression levels?

2. How does household income among those who are unemployed affect their mental health?

3. How does education level among those who are unemployed affect their mental health?

## Question 1:

How much does unemployment have an effect on stress, anxiety, and depression levels?

```{r}
# Create a new data frame with counts of "1" values for Stress, Anxiety, and Depression
count_data <- data3 %>%
  group_by(Employed) %>%
  summarise(
    across(c(Stress, Anxiety, Depression), ~ sum(!is.na(.) & . == 1)),
    .groups = 'drop'
  )

# Reshape the data into long format
count_data_long <- count_data %>%
  pivot_longer(cols = c(Stress, Anxiety, Depression), names_to = "Condition", values_to = "Count")

# Create the grouped bar graph
ggplot(count_data_long, aes(x = Condition, y = Count, fill = factor(Employed))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Effect of Unemployment on Anxiety, Depression, and Stress",
    x = "Condition",
    y = "Number of People with Condition"
  ) +
  scale_fill_manual(values = alpha(c("0" = "mistyrose3", "1" = "slategray3"), alpha = 0.7),  # alpha = transparency in color
                    breaks = c("0", "1"),
                    labels = c("Unemployed", "Employed"))  +
  labs(fill = "Employment Status") +
  theme_minimal() +
  theme(
        axis.title = element_text(size = 10, face = "bold"),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        legend.text = element_text(size = 8, face = "italic"),  # Make legend labels bold
        legend.title = element_text(size = 8)
  )
```

## Question 2:

How does household income among those who are unemployed affect their mental health?

```{r}
# Subset of data1 with household income included
data1b <- data1%>%
  select(I.am.currently.employed.at.least.part.time,I.identify.as.having.a.mental.illness,Household.Income)%>%
  mutate(Employed = I.am.currently.employed.at.least.part.time, Mental_Illness_Status = I.identify.as.having.a.mental.illness, Household_Income = Household.Income)
data1b <- data1b%>%select(Employed, Mental_Illness_Status,Household_Income)

# Extract only the unemployed
data1b_unemployed <- data1b%>%filter(Employed == 0)
data1b_unemployed$Mental_Illness_Status <- factor(data1b_unemployed$Mental_Illness_Status, labels = c("No Mental Illness", "Has Mental Illness"))

# Filter out "Prefer not to answer"
data1b_unemployed <- data1b_unemployed%>%filter(Household_Income != "Prefer not to answer")

# Plot the graph
ggplot(data1b_unemployed, aes(x = Household_Income, fill = Mental_Illness_Status)) +
  geom_bar(position = "fill") +
  scale_x_discrete(limits = c("$0-$9,999", "$10,000-$24,999","$25,000-$49,999","$50,000-$74,999","$75,000-$99,999", "$100,000-$124,999", "$200,000+")) +
  scale_fill_manual(values = alpha(c("No Mental Illness" = "slategray3", "Has Mental Illness" = "mistyrose3"), alpha = 0.7),
                    labels = c("No Mental Illness", "Has Mental Illness")) +
  labs(title = "Scaled Stacked Bar Chart of Household Income vs. Mental Illness Among Unemployed People",
       x = "Household Income",
       y = "Proportion of People",
       fill = "Mental Health Status") +
  theme_minimal() + 
  theme(
        title = element_text(size = 8),
        axis.text.x = element_text(size = 5),
        axis.title = element_text(size = 8, face = "bold"),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        legend.text = element_text(size = 8, face = "italic"),
        legend.title = element_text(size = 8)
  )
```

## Question 3: 

How does education level among those who are unemployed affect their mental health?

```{r}
data1a$Mental_Illness_Status <- factor(data1a$Mental_Illness_Status, labels = c("No Mental Illness", "Has Mental Illness"))

data1a_unemployed <- data1a %>% filter(Employed == 0)
#view(data1a_unemployed)
ggplot(data1a_unemployed, aes(x = Education, fill = Mental_Illness_Status)) +
  geom_bar(position = "fill") +
  scale_x_discrete(limits = c("Dropped Out", "High School", "Undergraduate", "Masters", "PHD")) +
  scale_fill_manual(values = alpha(c("No Mental Illness" = "slategray3", "Has Mental Illness" = "mistyrose3"), alpha = 0.7),
                    labels = c("No Mental Illness", "Has Mental Illness")) +
  labs(title = "Scaled Stacked Bar Chart of Education vs. Mental Illness Among Unemployed People",
       x = "Education Level",
       y = "Proportion of People",
       fill = "Mental Health Status") +
  theme_minimal() + 
  theme(
        axis.title = element_text(size = 10, face = "bold"),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        legend.text = element_text(size = 8, face = "italic"),
        legend.title = element_text(size = 8)
  )
```

## Conclusions

1. Unemployment had an obvious negative effect on factors related to mental health, as shown by the increased rates of anxiety, stress, and depression

2. Household income had no obvious correlation with mental health illness.

3. Having higher education did not appear to have a strong correlation between being unemployed and having a mental illness.


## Data Sources:

[Unemployment and Mental Illness Survey](https://www.kaggle.com/datasets/michaelacorley/unemployment-and-mental-illness-survey)

[Medical Student Mental Health](https://www.kaggle.com/datasets/thedevastator/medical-student-mental-health?select=Data+Carrard+et+al.+2022+MedTeach.csv)
